Don't Mind the Gap: Bridging Network-wide Objectives and Device-level Configurations

### 2016-06-19 paper/v2/paper-edited.tex
edits for sections 1-4

As a complementary approach, configuration analysis tools can help reduce misconfigurations by checking if low-level configurations match high-level policy~\cite{batfish,feamster+:rcc}. However, such tools too cannot help operators with the challenging task of generating configurations in the first place.

@@

To implement policy P1, the operators must compute and assign
local preferences such that preferences at \CD{Cust}-facing interfaces
$>$ \CD{Peer}-facing interfaces $>$ \CD{Prov}-facing interfaces. At
the same time, to satisfy P3, the preference at \CD{R2}'s
\CD{Cust}-facing interface should be lower than that at
\CD{R1}. Implementing P3 will also require MEDs to be appropriately configured on \CD{R1} and \CD{R2}.
To implement P2, the operators can assign communities that
indicate where a certain routing announcement entered the
network. Then, \CD{R4} must be configured to not announce to
$\CD{Peer}$ routes that have communities that correspond to the
\CD{R2}-\CD{Prov} link but to announce routes with communities for the \CD{R2}-\CD{Cust} and \CD{R1}-\CD{Cust} links. A similar policy must be configured for \CD{R2}. Finally, to implement P4 and P5, the operators will have to compute and configure appropriate prefix- and AS-path-based import and export filters at each router.

@@

However, this implementation has incorrect behavior in the face of failures. Suppose links X--G and X--H fail. Then, X will hear announcements for \CD{PL*} from C and D, having traversed from G and H to Y to C and D. Per its policy implementation, X will start "leaking" these prefixes externally. Depending on the rationale for local services, this leak could impact security (e.g., if the services are sensitive) or availability (e.g., if the \CD{PL*} prefixes are reused for other services outside of the datacenter). This problem does not manifest without failures because then X has and prefers paths to \CD{PL*} through G and H since they are shorter. A similar problem will occur if links Y--G and Y--H fail.
%\footnote{A different problem occurs when links X--C and X--D (or Y--C and Y--D) fail. X (Y) may stop announcing the global prefixes because they would be heard through G and H.}
Link failures in datacenters are frequent and it is not uncommon to have multiple failed links at a given time~\cite{dc-failure-study}.

To avoid this problem, the operator may disallow "valley" paths, i.e., those that go up, down, and back up again. This guard can be implemented by $X$ and $Y$ rejecting paths through the other. But that creates a different problem in the face of failures---an aggregation-induced blackhole~\cite{route-aggregation}. If links D--A and X--C fail, X will hear announcements for \CD{PG2} from D and will thus announce \CD{PG} externally. This announcement will bring to X traffic for \CD{PG1} as well, but because valleys are disallowed, X does not have a valid route for \CD{PG1} and will drop all traffic to it.

@@

\sysname lets operators configure the network with the
abstraction that they have centralized control over routing.
Specifically, the operator simply provides a set of high-level constraints
that describe the
paths traffic should---or should not---take and their relative preferences.
\sysname specifications are written modularly via a series
of declarations.
%These definitions allow users to express the three
%major elements of any \sysname specification:  \emph{prefixes},
%\emph{paths} and \emph{policies}.
For example, to begin specification of the backbone
network from the previous section, we first express the idea that
we prefer that traffic leave the network through \CD{R1} over \CD{R2} (to \CD{Cust}) over \CD{Peer} over \CD{Prov} (policy P3 from Figure~\ref{fig:example1}):
%\begin{lstlisting}[mathescape=true]
%define ExitCust = $exit(R_1 \gg R_2)$
%\end{lstlisting}
\begin{code}
\Define Prefs = \Exit(R1 \Prefer R2 \Prefer Peer \Prefer Prov)
\end{code}
This  statement defines a set of \emph{ranked paths}, which includes
all paths (and only those paths) for which traffic exits our network
through either router \CD{R1}, router \CD{R2}, \CD{Peer}, or \CD{Prov}.
The paths that exit through \CD{R1}
are preferred (\Prefer) to those that exit through \CD{R2}, which are preferred to those that
leave through \CD{Peer} and then \CD{Prov}.  As we
describe in the next section, the \Exit\ predicate, as well as other
path predicates used later in this section, is simply a shorthand for
a particular regular expression over paths that is expressible in our policy
language. The preference operator (\Prefer) is flexible and can be used between constraints as well as among individual routers. For example, the above constraint could have been written equivalently as
\Exit(\CD{R1}) \Prefer \ldots \Prefer ~\Exit(\CD{Prov})

@@

Since, routing allows transit traffic by default (\EG, traffic entering from
\CD{Peer} and leaving through \CD{Prov}), we separately define a
policy to enforce P2 and P5 from Figure~\ref{fig:example1}, using conjunction (\AND),
disjunction (\OR) and negation (\NOT) of constraints.
First, we create reusable abstractions for describing traffic that transits our network. In \sysname, this is done by creating a new parameterized definition.
\begin{code}
\Define transit(X,Y) = \Enter(X{\OR}Y) \AND \Exit(X{\OR}Y)
\Define cust-transit(X,Y) = \Later(X) \AND \Later(Y)
\end{code}
\noindent
Here we define transit traffic between groups of neighbors $X$ and $Y$ as traffic that enters the network through some neighbor in $X$ or $Y$ and then leaves the network through some neighbor in either $X$ or $Y$. Similarly, we define customer transit for customer $X$ and provider $Y$ as traffic that later goes through both $X$ and $Y$ after leaving our network. Using these two new abstractions, we can now implement policies P2 and P5 with the following constraint.

@@

\begin{code}
\Define Ownership =
    \{PG1 \Path \End(A)
     PG2 \Path \End(B)
     PL1 \Path \End(E)
     PL2 \Path \End(F)
     \True \Path \End(\Out)\}
\end{code}
\noindent
This definition says that traffic for prefix \CD{PG1} is allowed to follow only paths that
end at router \CD{A}; traffic for \CD{PG2},
but not \CD{PG1}, must
end at router \CD{B}; and so on.
Any traffic destined for a prefix that not a part of the datacenter should be allowed to leave the datacenter and end at some external location, which is otherwise unconstrained.
The special keyword \Out{} matches any location outside the datacenter network, while the keyword \In{} will match any location inside the network.

@@

%\para{Syntax}
The \sysname FE is just a thin layer atop the RIR for describing preference-based path constraints.
Figure~\ref{fig:rir-syntax} shows the RIR syntax. A policy has one or more constraints, each of which has a test on the type of route, and a corresponding set of preferred regular paths. Regular paths are regular expressions where the base characters are abstract locations---either a router or an AS. Special \In{} and \Out{} symbols refer to any internal or external location respectively. In addition, $\Sigma$ refers to any symbol. We also use the standard regular expressions abbreviation $r^+$ for $r \cdot r^*$, a sequence of one or more occurrences of $r$. Predicates ($t$) consist of logical boolean connectives (and, or, not) as well as tests that match a particular prefix (or group of prefixes) and tests for route advertisments with a particular community value (i.e., an integer value associated with a path).
%% For example, the constraint
%% \[(\textit{prefix}=74.3.28.0/24) \Link \\
%% (as200 \cdot \In^+) \Prefer (as100 \cdot \In^+)
%% \]
%% describes a more-preferred set of paths for traffic announced by a prefix no less specific than \CD{74.125.28.0/24}, which starts at AS 200, before entering and staying inside the user's network to get to the destination, and a less-preferred set of paths that start at AS 100 and are otherwise the same. The plus operator $\In^+$ stands for $\In \cdot \In^*$, at least one internal hop. Tests over route types use standard boolean connectives, and can refer to both prefixes and route community values.

@@

The translation from FE to RIR is based on a set of rewriting rules.
The first step merges separate constraints. It takes the cross product of per-prefix constraints, where logical conjunction ($a \AND b$) is replaced by intersection on regular constraints ($a \Intersect b$), logical disjunction is replaced by union, and logical negation ($\NOT a$) is replaced by path negation ($\Any \cap \NOT(a)$). The additional constraint $\Any$ ensures the routes are well-formed by restricting the paths to only those that go through the user's network.
%
For example, in the datacenter configuration from \S\ref{sec:propane}, combining the \CD{Locality} and \CD{Ownership} policies results in the following RIR:

@@

The next step rewrites the high-level constraints such as $\Enter$ according to the equivalences in Figure~\ref{fig:rir-syntax}. Since preferences can only occur at the outermost level for an RIR expression, the final step is to ``lift" occurrences of the preference operator in each regular expression. For example, the regular expression $a \cdot (b \Prefer c) \cdot d$ is lifted to $(a \cdot b \cdot d) \Prefer (a \cdot c \cdot d)$ by distributing the preference over the sequence operator. In general, we employ the follwing distributivity equivalences:
%
\[
\begin{array}{c}
  x \odot (y_1 \Prefer \dots \Prefer y_n) = (x \odot y_1) \Prefer \dots \Prefer (x \odot y_n) \\
  (y_1 \Prefer \dots \Prefer y_n) \odot x = (y_1 \odot x) \Prefer \dots \Prefer (y_n \odot x)
\end{array}
\]
%
where $\odot$ stands for an arbitrary regular binary operator, and $x$ is a policy with a single preference. The compiler flags as invalid preferences nested under a unary operator (i.e., \textit{star} or \textit{negation}).

@@

\begin{figure*}
  \begin{minipage}[t]{.5\linewidth}

@@

  \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=1.6cm]
    \state{0}{$0$}{              }{}
    \state{1}{$1$}{right of=0}{}
    \state{2}{$2$}{right of=1}{}
    \state{3}{$3$}{right of=2}{}
    \state{4}{$4$}{right of=3}{accepting}
    \transition{0}{1}{out}{}
    \transition{1}{2}{in}{}
    \transition{2}{2}{A,C,D,E}{loop above}
    \transition{2}{3}{B}{}
    \transition{3}{3}{B}{loop above}
    \transition{3}{2}{A,C,D,E}{bend left}
    \transition{3}{4}{W}{}
  \end{tikzpicture}
  \end{minipage}
  %
  ~~
  ~~
  %
  \begin{minipage}[t]{.5\linewidth}
  \hdr{Product Graph IR}{}
  \vspace*{-1\baselineskip}
  %
  \includegraphics[width=.65\columnwidth]{figures/productgraph}
  \end{minipage}

@@

% Formally define automata
%The RIR policy is an ordered sequence of regular expressions $r_1 \Prefer \dots \Prefer r_j$.
While paths in RIR policy describe the direction traffic flow through the network, to implement the policy with BGP we are concerned about the way control-plane information is disseminated, i.e., route announcements flowing in the opposite direction. To capture this idea, for each regular expression $r_i$ in RIR policy, we construct a deterministic finite state machine on the reversed regular expression. Each automaton for the reversed regular expression of $r_i$ is a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ consists of all abstraction locations (i.e., routers or ASes), $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
% Formally define topology
The topology is represented as a graph ($V, E$), which consists of a set of vertices $V$, and a set of directed edges $E \colon V \times V$.
%
% Formally define product graph
The combined PGIR is a tuple ($V'$, $E'$, $s$, $e$, $P$) with
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $s$,
a unique ending vertex $e$,
and a preference function $P \colon V' \rightarrow 2^{\set{1, \dots, j}}$ , which maps nodes in the product graph to a set of preference ranks.
For a PGIR vertex $v' = (v, \dots) \in V'$, we say $loc(v') = v$, which indicates that the corresponding topology location for $v'$ is $v$.
Two nodes $x$ and $y$ \emph{shadow} each other in the product graph when $x \in V'$ and $y \in V'$ and $loc(x) = loc(y)$.

@@

% Formal product construction
Let $a_i$ and $b_i$ denote states in the regular policy automata.
The PGIR is constructed by adding an edge from $v_1 = (x, a_1, \dots, a_m)$ to $v_2 = (y, b_1, \dots, b_m)$ whenever $\sigma_i(a_i, y) = b_i$ for each $i$ and $(x,y) \in E$ is a valid topology link.
%
Additionally, we add edges from the start node $s$ to any $v = (x, a_1, \dots, a_m)$ when $\sigma_i(q_{0_i}, x) = a_i$ for each $i$.
%
The preference function $P$ for node $v = (x, a_1, \dots, a_m)$ is defined as $P(v) = \set{i~\vert~a_i \in F_i} $. That is, it records which preferences are achieved from each policy automaton for every node in the product construction.
%
Finally, there is an edge from each node in the PGIR such that $P(v) \neq \emptyset$ to the special end node $e$. We write $(x \leq_{rank} y)$ if either $P(x) = P(y) = \emptyset$ or $min ~ P(x) \leq min ~ P(y)$, which means that paths ending at PGIR node $x$ are better than (lower rank) than paths that ending at $y$.

@@

We use graph dominators~\cite{tarjan-dominance} as a cheap approximation for removing many nodes and edges in the PGIR that are never on any \emph{simple} (loop free) path between the start and end nodes.
In the PGIR, a node $x$ dominates a node $y$ if $x$ appears on every path leading from the start node to $y$. Similarly, a node $x$ post-dominates a node $y$ in the PGIR if $x$ appears on every path from $y$ to the end node.
We can safely remove nodes and edges in the PGIR when any of the following conditions hold, where $x$, $x'$ and $y$, $y'$ refer to PGIR nodes that shadow each other:
\begin{itemize}
\setlength{\itemsep}{1pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}
\item Remove $x$ if it is not reachable from the start node
\item Remove $x$ if it can not reach the end node
\item Remove $x$ if it is (post-)dominated by some $x'$
\item Remove edge ($x$, $y$) if $y$ is post-dominated by some $x'$
\item Remove edge ($x$, $y$) if $x$ is dominated by some $y'$
\end{itemize}
%We can safely remove any PGIR node or edge as long as it never appears on any \emph{simple} path from the \textit{start} node to the \emph{end} node.
For example, node $(W,1,1)$ in Figure~\ref{fig:example-compilation} can be removed since it is never on a simple path to the end node because it must always go through node $(W,-,4)$. That is, node $(W,-,4)$ post-dominates node $(W,1,1)$. We repeatedly apply the minimizations above to the PGIR until no further minimization is possible.
In the example of Figure~\ref{fig:example-compilation}, colored nodes and dashed edges are removed after minimization.
 %since they are irrelvant to the BGP decision process.
%
%Although fully minimizing this graph is an NP-complete problem,  a simple and efficient algorithm based on graph dominators achieves largely the same effect, greatly simplifying the failure safety analysis.

%A node $x$ dominates $y$ if $x$ appears on every path leading to $y$ from a source node. Many efficient algorithms exist for finding graph dominators~\ref{bib:todo}. For minimization, we compute the dominator set for each node in both the product graph $G$ with respect to the start node, and in the reversed product graph $G^R$ with respect to the end node. The following rules enable efficient minimization of the product graph.
%
%\begin{itemize}
%  \item Remove nodes that never reach the end node
%  \item Remove nodes not reachable from the $start$ node
%  \item Remove any node $x$ such that $x$ is dominated by $y$ in $G$ or $G^R$, and $loc(x) = loc(y)$
%  \item Remove any edge from $x$ to $y$ in $G$ or $G^R$ if there is some node $z$ dominated by $y$ such that $loc(x) = loc(z)$
%\end{itemize}
%
%Repeated application of the above rules leads to removing the colored nodes and dashed lines in Figure~\ref{fig:example-compilation}.

@@

After configuration generation, the compiler further processes the ABGP policy, removing community tags when possible, combining filters, removing dead filters, and so on. In the compilation example, all community tags can be removed since there is never any ambiguity based on the router importing the route, and the neighbor from which the route is being imported. Similarly, after removing the communities, the two rules at router $B$ are merged into a single rule.

@@

Constructing automata for extended regular expressions (i.e., those that include negation and intersection) are known to have high complexity~\cite{regex-complexity}. The \sysname compiler uses regular expressions derivatives~\cite{regex-derivatives} with character classes to construct deterministic automata for extended regular expressions over large alphabets efficiently. Since regular expressions are defined over a finite alphabet, and since much of the AS topology is unknown, we set the alphabet to include all uniquely referenced external ASes in the policy. Similarly, to model the unknown external AS topology beyond immediate peers, we include a special topology node to represent any unknown location.
%
Rather than construct the product graph in full, our implementation prevents exploring parts of the graph when no accepting state is reachable in any of the corresponding regular automata during construction.

>>>>

As a complementary approach, configuration analysis tools can help reduce misconfigurations by checking if low-level configurations match high-level policy~\cite{batfish,feamster+:rcc}. However, such tools cannot help operators with the challenging task of generating configurations in the first place.

@@

To implement policy P1, the operators must compute and assign
local preferences such that preferences at \CD{Cust}-facing interfaces
$>$ \CD{Peer}-facing interfaces $>$ \CD{Prov}-facing interfaces. At
the same time, to satisfy P3, the preference at \CD{R2}'s
\CD{Cust}-facing interface should be lower than that at
\CD{R1}. Implementing P3 will also require MEDs to be appropriately configured on \CD{R1} and \CD{R2}.
To implement P2, the operators can assign communities that
indicate where a certain routing announcement entered the
network. Then, \CD{R4} must be configured to not announce to
$\CD{Peer}$ routes that have communities that correspond to the
\CD{R2}-\CD{Prov} link but to announce routes with communities for the \CD{R2}-\CD{Cust} and \CD{R1}-\CD{Cust} links. A similar policy must be configured for \CD{R2} as well. Finally, to implement P4 and P5, the operators will have to compute and configure appropriate prefix- and AS-path-based import and export filters at each router.

@@

However, this implementation has incorrect behavior in the face of failures. Suppose links X--G and X--H fail. Then, X will hear announcements for \CD{PL*} from C and D, having traversed from G and H to Y to C and D. Per its policy implementation, X will start "leaking" these prefixes externally. Depending on the rationale for local services, this leak could impact security (e.g., if the services are sensitive) or availability (e.g., if the \CD{PL*} prefixes are reused for other services outside of the datacenter). This problem does not manifest without failures because then X has and prefers paths to \CD{PL*} through G and H since they are shorter. A similar problem will occur if links Y--G and Y--H fail.
%\footnote{A different problem occurs when links X--C and X--D (or Y--C and Y--D) fail. X (Y) may stop announcing the global prefixes because they would be heard through G and H.}
Link failures in datacenters are frequent and it is not uncommon to have many failed links at a given time~\cite{dc-failure-study}.

To avoid this problem, the operator may disallow "valley" paths, i.e., those that go up, down, and back up again. This guard can be implemented by $X$ and $Y$ rejecting paths through the other. But that creates a different problem in the face of failures---an aggregation-induced blackhole~\cite{route-aggregation}. If links D--A and X--C fail, X will hear an announcement for \CD{PG2} from D and will thus announce \CD{PG} externally. This announcement will bring to X traffic for \CD{PG1} as well, but because valleys are disallowed, X does not have a valid route for \CD{PG1} and will drop all traffic for it despite the fact that a valid path exists through $Y$.

@@

\sysname lets operators configure the network with the
abstraction that they have centralized control over routing.
Specifically, the operator simply provides a set of high-level constraints
that describe the
paths traffic should---or should not---take and their relative preferences.
\sysname specifications are written modularly via a series
of declarations.
%These definitions allow users to express the three
%major elements of any \sysname specification:  \emph{prefixes},
%\emph{paths} and \emph{policies}.
For example, to begin specification of the backbone
network from the previous section, we first express the idea that
we prefer that traffic leave the network through \CD{R1} over \CD{R2} (to \CD{Cust}) over \CD{Peer} over \CD{Prov} (policy P1 and P3 from Figure~\ref{fig:example1}):
%\begin{lstlisting}[mathescape=true]
%define ExitCust = $exit(R_1 \gg R_2)$
%\end{lstlisting}
\begin{code}
\Define Prefs = \Exit(R1 \Prefer R2 \Prefer Peer \Prefer Prov)
\end{code}
This  statement defines a set of \emph{ranked paths}, which includes
all paths (and only those paths) for which traffic exits our network
through either router \CD{R1}, router \CD{R2}, \CD{Peer}, or \CD{Prov}.
The paths that exit through \CD{R1}
are preferred (\Prefer) to those that exit through \CD{R2}, which are preferred to those that
leave through \CD{Peer} and then \CD{Prov}.  As we
describe in the next section, the \Exit\ predicate, as well as other
path predicates used later in this section, are simply a shorthand for
a particular regular expression over paths that is expressible in our policy
language. The preference operator (\Prefer) is flexible and can be used between constraints as well as among individual routers. For example, the above constraint could have been written equivalently as
\Exit(\CD{R1}) \Prefer \ldots \Prefer ~\Exit(\CD{Prov})

@@

Since, routing allows transit traffic by default (\EG, traffic entering from
\CD{Peer} and leaving through \CD{Prov}), we separately define a
policy to enforce P2 and P5 from Figure~\ref{fig:example1}, using conjunction (\AND),
disjunction (\OR) and negation (\NOT) of constraints.
First, we create reusable abstractions for describing traffic that transits our network. In \sysname, this is done by creating a new parameterized definition.
\begin{code}
\Define transit(X,Y) = \Enter(X{\OR}Y) \AND \Exit(X{\OR}Y)
\Define cust-transit(X,Y) = \Later(X) \AND \Later(Y)
\end{code}
\noindent
Here we define transit traffic between groups of neighbors $X$ and $Y$ as traffic that enters the network through some neighbor in $X$ or $Y$ and then also leaves the network through some neighbor in either $X$ or $Y$. Similarly, we define customer transit for customer $X$ and provider $Y$ as traffic that later goes through both $X$ and $Y$ after leaving our network. Using these two new abstractions, we can now implement policies P2 and P5 with the following constraint.

@@

\begin{code}
\Define Ownership =
    \{PG1 \Path \End(A)
     PG2 \Path \End(B)
     PL1 \Path \End(E)
     PL2 \Path \End(F)
     \True \Path \End(\Out)\}
\end{code}
\noindent
This definition says that traffic for prefix \CD{PG1} is allowed to follow only paths that
end at router \CD{A}; traffic for \CD{PG2},
but not \CD{PG1}, must
end at router \CD{B}; and so on.
Any traffic destined for a prefix that is not a part of the datacenter should be allowed to leave the datacenter and end at some external location, which is otherwise unconstrained.
The special keyword \Out{} matches any location outside the datacenter network, while the keyword \In{} will match any location inside the network.

@@

%\para{Syntax}
The \sysname FE is just a thin layer atop the RIR for describing preference-based path constraints.
Figure~\ref{fig:rir-syntax} shows the RIR syntax. A policy has one or more constraints, each of which has a test on the type of route, and a corresponding set of preferred regular paths. Regular paths are regular expressions where the base characters are abstract locations---either a router or an AS. Special \In{} and \Out{} symbols refer to any internal or external location respectively. In addition, $\Sigma$ refers to any symbol. We also use the standard regular expression abbreviation $r^+$ for $r \cdot r^*$, a sequence of one or more occurrences of $r$. Predicates ($t$) consist of logical boolean connectives (and, or, not) as well as tests that match a particular prefix (or group of prefixes) and tests for route advertisments with a particular community value (i.e., an integer value associated with a path).
%% For example, the constraint
%% \[(\textit{prefix}=74.3.28.0/24) \Link \\
%% (as200 \cdot \In^+) \Prefer (as100 \cdot \In^+)
%% \]
%% describes a more-preferred set of paths for traffic announced by a prefix no less specific than \CD{74.125.28.0/24}, which starts at AS 200, before entering and staying inside the user's network to get to the destination, and a less-preferred set of paths that start at AS 100 and are otherwise the same. The plus operator $\In^+$ stands for $\In \cdot \In^*$, at least one internal hop. Tests over route types use standard boolean connectives, and can refer to both prefixes and route community values.

@@

The translation from FE to RIR is based on a set of rewriting rules.
The first step merges separate constraints. It takes the cross product of per-prefix constraints, where logical conjunction ($a \AND b$) is replaced by intersection on regular constraints ($a \Intersect b$), logical disjunction is replaced by union, and logical negation ($\NOT a$) is replaced by path negation ($\Any \cap \NOT(a)$). The additional constraint $\Any$ ensures the routes are well-formed by restricting the paths to only those that go through the user's network.
%
For example, in the datacenter FE configuration from \S\ref{sec:propane}, combining the \CD{Locality} and \CD{Ownership} policies results in the following RIR:

@@

The next step rewrites the high-level constraints such as $\Enter$ according to the equivalences in Figure~\ref{fig:rir-syntax}. Since preferences can only occur at the outermost level for an RIR expression, the final step is to ``lift" occurrences of the preference operator in each regular expression. For example, the regular expression $a \cdot (b \Prefer c) \cdot d$ is lifted to $(a \cdot b \cdot d) \Prefer (a \cdot c \cdot d)$ by distributing the preference over the sequence operator. In general, we employ the follwing distributivity equivalences:
%
\[
\begin{array}{c}
  x \odot (y_1 \Prefer \dots \Prefer y_n) = (x \odot y_1) \Prefer \dots \Prefer (x \odot y_n) \\
  (y_1 \Prefer \dots \Prefer y_n) \odot x = (y_1 \odot x) \Prefer \dots \Prefer (y_n \odot x)
\end{array}
\]
%
where $\odot$ stands for an arbitrary regular binary operator, and $x$ is a policy with a single preference. The compiler flags preferences nested under a unary operator (i.e., \textit{star} or \textit{negation}) as invalid.

@@

\begin{figure*}[t!]
  \begin{minipage}[t]{.5\linewidth}

@@

  \vspace*{0.5em}

  \begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=1.6cm]
    \state{0}{$0$}{              }{}
    \state{1}{$1$}{right of=0}{}
    \state{2}{$2$}{right of=1}{}
    \state{3}{$3$}{right of=2}{}
    \state{4}{$4$}{right of=3}{accepting}
    \transition{0}{1}{out}{}
    \transition{1}{2}{in}{}
    \transition{2}{2}{A,C,D,E}{loop above}
    \transition{2}{3}{B}{}
    \transition{3}{3}{B}{loop above}
    \transition{3}{2}{A,C,D,E}{bend left}
    \transition{3}{4}{W}{}
  \end{tikzpicture}
  \end{minipage}
  %
  ~~
  ~~
  %
  \begin{minipage}[t]{.5\linewidth}
  \hdr{Product Graph IR}{}
  \vspace*{-1\baselineskip}
  %
  \includegraphics[width=.8\columnwidth]{figures/productgraph}
  \end{minipage}

@@

% Formally define automata
%The RIR policy is an ordered sequence of regular expressions $r_1 \Prefer \dots \Prefer r_j$.
While paths in RIR policy describe the direction traffic flow through the network, to implement the policy with BGP we are concerned about the way control-plane information is disseminated, i.e., route announcements flowing in the opposite direction. To capture this idea, for each regular expression $r_i$ in RIR policy, we construct a deterministic finite state machine on the reversed regular expression. Each automaton for the reversed regular expression of $r_i$ is a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ consists of all abstraction locations (i.e., routers or ASes), $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
% Formally define topology
The topology is represented as a graph ($V, E$), which consists of a set of vertices $V$, and a set of directed edges $E \colon V \times V$.
%
% Formally define product graph
The combined PGIR is a tuple ($V'$, $E'$, $s$, $e$, $P$) with
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $s$,
a unique ending vertex $e$,
and a preference function $P \colon V' \rightarrow 2^{\set{1, \dots, j}}$ , which maps nodes in the product graph to a set of preference ranks.
For a PGIR vertex $v_1 = (v, \dots) \in V'$, we say $loc(v_1) = v$, which indicates that the corresponding topology location for $v_1$ is $v$.

%
We say that a PGIR node $x_1$ is a \emph{shadow} of topology location $x$ when $x_1 \in V'$ and $loc(x_1) = x$.
%
%Throughout the remainder of the paper, we will use metavariables without subscripts such as $x$ to represent topology locations, and we will use similar metavariables with subscripts such as $x_1$ or $x_2$ to represent shadows of $x$ (i.e., PGIR nodes for location $x$).

@@

% Formal product construction
Let $a_i$ and $b_i$ denote states in the regular policy automata.
The PGIR is constructed by adding an edge from $v_1 = (x, a_1, \dots, a_m)$ to $v_2 = (y, b_1, \dots, b_m)$ whenever $\sigma_i(a_i, y) = b_i$ for each $i$ and $(x,y) \in E$ is a valid topology link.
%
Additionally, we add edges from the start node $s$ to any $v = (x, a_1, \dots, a_m)$ when $\sigma_i(q_{0_i}, x) = a_i$ for each $i$.
%
The preference function $P$ for node $v = (x, a_1, \dots, a_m)$ is defined as $P(v) = \set{i~\vert~a_i \in F_i} $. That is, it records which preferences are achieved from each policy automaton for every node in the product construction.
%
Finally, there is an edge from each node in the PGIR such that $P(v) \neq \emptyset$ to the special end node $e$. We write $(x \leq_{rank} y)$ if either $P(x) = P(y) = \emptyset$ or $min ~ P(x) \leq min ~ P(y)$, which means that paths ending at PGIR node $x$ are better (lower rank) than paths ending at $y$.

@@

We use graph dominators~\cite{tarjan-dominance} as a cheap approximation for removing many nodes and edges in the PGIR that are never on any \emph{simple} (loop free) path between the start and end nodes.
In the PGIR, a node $x$ dominates a node $y$ if $x$ appears on every path leading from the start node to $y$. Similarly, a node $x$ post-dominates a node $y$ in the PGIR if $x$ appears on every path from $y$ to the end node.
We can safely remove nodes and edges in the PGIR when any of the following conditions hold, where $x$, $x'$ and $y$, $y'$ refer to PGIR nodes that are shadows of the same topology location.
\begin{itemize}
\setlength{\itemsep}{1pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}
\item Remove $x$ if it is not reachable from the start node
\item Remove $x$ if it can not reach the end node
\item Remove $x$ if it is (post-)dominated by some $x'$
\item Remove edge ($x$, $y$) if $y$ is post-dominated by some $x'$
\item Remove edge ($x$, $y$) if $x$ is dominated by some $y'$
\end{itemize}
For example, node $(W,1,1)$ in Figure~\ref{fig:example-compilation} is removed since it is never on a simple path to the end node because the path must always go through node $(W,-,4)$. That is, node $(W,-,4)$ post-dominates node $(W,1,1)$ and both shadow topology location $W$. 
Similarly, the edge from $(C,3,2)$ to $(D,-,2)$ is removed since node $(C,-,2)$ post-dominates $(D,-,2)$ and shadows the same topology location $C$. 
We repeatedly apply the minimizations above until no further minimization is possible.
In the example of Figure~\ref{fig:example-compilation}, colored nodes and dashed edges are removed after minimization.

@@

After configuration generation, the compiler further processes the ABGP policy, removing community tags when possible, combining filters, removing dead filters, and so on. In the compilation example, all community tags can be removed since there is never any ambiguity based on the router importing the route, and the neigbor the route is being imported from. Similarly, after removing the communities, the two rules at router $B$ are merged into a single rule.

@@

Constructing automata for extended regular expressions (i.e., regular expressions with negation and intersection operations) is known to have high complexity~\cite{regex-complexity}. The \sysname compiler uses regular expressions derivatives~\cite{regex-derivatives} with character classes to construct deterministic automata for extended regular expressions over large alphabets efficiently. Since regular expressions are defined over a finite alphabet, and since much of the AS topology is unknown, we set the alphabet to include all uniquely referenced external ASes in the policy. Similarly, to model the unknown external AS topology beyond immediate peers, we include a special topology node to represent any unknown location.
%
Rather than construct the product graph in full, our implementation prevents exploring parts of the graph when no accepting state is reachable in any of the corresponding regular automata during construction.
